---
title: "Pan-Activity in ToxCast/Tox21"
author: "Kimberly Truong"
date: '2023-09-26'
output:
  html_document:
    code_folding: hide
    collapsed: yes
    df_print: paged
    lightbox: no
    number_sections: no
    self_contained: yes
    thumbnails: no
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

# Data Import and Environment

```{r setup, include = FALSE}
library(knitr)
knitr::opts_knit$set(root.dir = 'C:/Users/ktruong/OneDrive - Environmental Protection Agency (EPA)/Profile/Documents/thyroid_prioritization')

```

```{r, warning = FALSE, message = FALSE}
rm(list=ls())
library(tcpl)
library(RMySQL)
library(data.table)
library(dplyr)
library(ggplot2)
library(ggpubr) 

# loads in mc5/sc2 data after some coarse filtering
# long_hitc_final is a long df where every row is a chem x aeid pair and its sc/mc hitcalls
#' @detection_technology_type describes the assay tech
#' @cell_format 
load('./data/invitrodb_v3_5_snapshot.RData')

```

# Pull down ToxCast data {.tabset .tabset-fade .tabset-pills}
```{r, eval = FALSE}
# connect to invitrodb with tcpl
tcplConf(user='_dataminer', pass='pass', db='prod_internal_invitrodb_v3_5', drvr='MySQL', host='ccte-mysql-res.epa.gov') 

# pull down mc data 
mc5 <- tcplPrepOtpt(tcplSubsetChid(tcplLoadData(lvl=5, type = 'mc')))
mc6 <- tcplPrepOtpt(tcplLoadData(lvl=6, fld='m4id', val=mc5$m4id, type='mc'))
setDT(mc6)
mc6_mthds <- mc6[ , .( mc6_mthd_id = paste(mc6_mthd_id, collapse=",")), by = m4id]
mc6_flags <- mc6[ , .( flag = paste(flag, collapse=";")), by = m4id]
mc5$mc6_flags <- mc6_mthds$mc6_mthd_id[match(mc5$m4id, mc6_mthds$m4id)]
mc5[, flag.length := ifelse(!is.na(mc6_flags), count.fields(textConnection(mc6_flags), sep =','), NA)]

# filter the dataset, with coarse filters
mc5[hitc==1 & flag.length < 3, use.me := 1]
mc5[hitc==1 & is.na(flag.length), use.me := 1]
mc5[hitc==1 & flag.length >= 3, use.me := 0]
mc5[fitc %in% c(36,45), use.me := 0]
mc5[hitc==-1, use.me := 0] # data could not be curve-fitted (usually due to having < 4 concs)
mc5[use.me==0, modl_ga := as.numeric(NA)] 
mc5[use.me==0, hitc := 0] 
mc5[hitc==0, modl_ga := as.numeric(NA)]

setnames(mc5, c('dsstox_substance_id.x'), c('dsstox_substance_id'))
mc5[, dsstox_substance_id.y := NULL]

# pull down sc data 
sc2 <- tcplPrepOtpt(tcplSubsetChid(tcplLoadData(lvl=2, type='sc'), type = 'sc')) # for max response and hitcally summary
setnames(sc2, c('dsstox_substance_id.x'), c('dsstox_substance_id'))
sc2[, dsstox_substance_id.y := NULL]

# get hitcall data from sc/mc 
long.mc5 <- mc5[!is.na(dsstox_substance_id), 
                .(dsstox_substance_id, chnm, aeid, aenm, hitc)]
long.sc2 <- sc2[!is.na(dsstox_substance_id),  
                .(dsstox_substance_id, chnm, aeid, aenm, hitc)]

# combine mc/sc data
long_all <- merge.data.table(long.sc2, long.mc5, 
                             by = c("dsstox_substance_id", "chnm", "aeid", "aenm"), 
                             all = T)
setnames(long_all, old = c("hitc.x", "hitc.y"), 
         new = c("sc.hitc", "mc.hitc"))

# if tested in both, take hitc from mc; otherwise, take hitc from sc 
long_hitc_final <- long_all %>% 
  rowwise() %>%
  mutate(hitc = ifelse(!is.na(mc.hitc), mc.hitc, sc.hitc))

setDT(long_hitc_final)

```

```{r}
length(unique(long_hitc_final$aenm)) # 2228 unique endpoints in ToxCast/Tox21
length(unique(long_hitc_final$dsstox_substance_id)) # 9541 unique chems 
```
* 9541 unique chemicals
* 2228 total unique endpoints 

# Are colorful substances pan-actors? {.tabset .tabset-fade .tabset-pills}

## compute hit rates
```{r}
# by chemical hit-rate
# lw(x) == length(which(x))
hit.rates <- long_hitc_final[, list(
  total.cnt  = .N, #total number of aeids tested for that chem (sc and/or mc)
  active.cnt  = as.double(lw(hitc==1)),  # active count
  inactive.cnt  = as.double(lw(hitc==0)),  #inactive count
  active.pct = (lw(hitc==1)/.N)*100, #active percent
  inactive.pct = (lw(hitc==0)/.N)*100 #inactive percent
), by = list(chnm, dsstox_substance_id)]

# capture all colored chems 
color.chnms <- hit.rates$chnm[grep("\\w+\\s(red|orange|yellow|green|blue|violet|black)\\s*\\w*", hit.rates$chnm, ignore.case = T)]
ci <- hit.rates$chnm[grep("C.I.", hit.rates$chnm)] # important to capture  substances like "C.I. Azoic Diazo Component 112"
fdc <- hit.rates$chnm[grep("^[F]*D&C", hit.rates$chnm)]
all.colors.nm <- purrr::reduce(list(color.chnms, ci, fdc), union) # 97 chems 

hit.rates.colors <- hit.rates[chnm %in% all.colors.nm, ]

ggplot(hit.rates.colors, aes(x=active.pct)) + 
  geom_histogram(color="black", fill="#9DC183", bins = 15) + 
  geom_vline(aes(xintercept=median(hit.rates$active.pct)),
             color="blue", linetype="dashed", linewidth=1) +
  annotate(geom = "text", x = median(hit.rates$active.pct) + 3, y = 20, 
           label = paste0(round(median(hit.rates$active.pct), 1), "%")) +
  ggtitle('Hit Rates across all ToxCast Assays for Colorful Substances') +
  xlab('% of Assays found to be Active in') + ylab('Number of Chemicals') +
  theme_bw()

```

## what chemicals hit at least 30% of ToxCast assays?
```{r}
hit.rates.colors[active.pct >= 30][order(-active.pct)]

length(unique(hit.rates.colors[active.pct >= 30]$dsstox_substance_id))

```

## pull down assay meta data 
```{r}
# connect to invitrodb for RMySQL connection
con <- dbConnect(drv = RMySQL::MySQL(), user="ktruong", #fill user
                 password = "Today061822K!!", # fill pass
                 host = "ccte-mysql-res.epa.gov", database = prod_internal_invitrodb_v3_5)

assays <- dbGetQuery(con, "SELECT * FROM prod_internal_invitrodb_v3_5.assay 
INNER JOIN prod_internal_invitrodb_v3_5.assay_component ON assay.aid=assay_component.aid 
INNER JOIN prod_internal_invitrodb_v3_5.assay_component_endpoint ON assay_component.acid=assay_component_endpoint.acid;")

long_hitc_final$detection_technology_type <- assays$detection_technology_type[match(long_hitc_final$aeid, assays$aeid)]

```

## looking at just spectrophotometric assays
```{r}
sp.assays <- long_hitc_final[detection_technology_type %in% c("Spectrophotometry", "spectrophotometry", 
                                                  "Absorbance ", "Colorimetric")]

length(unique(sp.assays$aeid)) # 93 relevant colorimetric endpoints 

```

```{r}
hitrates.sp <- sp.assays[, list(
  total.cnt  = .N, #total number of aeids tested for each chem
  active.cnt  = as.double(lw(hitc==1)),  # active count
  inactive.cnt  = as.double(lw(hitc==0)),  #inactive count
  active.pct = (lw(hitc==1)/.N)*100, #active percent
  inactive.pct = (lw(hitc==0)/.N)*100 #inactive percent
), by = list(chnm, dsstox_substance_id)]

colors.data <- hitrates.sp[chnm %in% all.colors.nm, ]

ggplot(colors.data, aes(x=active.pct)) + 
  geom_histogram(color="black", fill="#00A86B", bins = 10) + 
  geom_vline(aes(xintercept=median(hitrates.sp$active.pct)),
             color="blue", linetype="dashed", linewidth=1) +
  annotate(geom = "text", x = median(hitrates.sp$active.pct) + 4, y = 17.5, 
           label = paste0(median(hitrates.sp$active.pct), "%")) +
  ylim(0, 20) +
  ggtitle('Hit Rates across all Spectrophotometric Assays for Colorful Substances') +
  xlab('% of Assays found to be Active in') + ylab('Number of Chemicals') +
  theme_bw()

```

## what are the chemicals that hit 100% of spectrophotometric assays?
```{r}
colors.data[active.pct == 100]

```

* FD&C Green No. 3 and Bromophenol Blue were found to be "active" for DIO1,2,3 and IYD targets. 
* C.I. Direct Yellow 12 was found to be active for DIO 1-3. 
* C.I. Acid Red 114 was only tested for IYD and was found to be active. 

## what are the chemicals that hit at least 30% of spectrophotometric assays? 
```{r}
colors.data[active.pct > 30 & active.pct <= 100]

```

## what are the chemicals that do not hit any of the spectrophotometric assays?
```{r}
colors.data[active.pct == 0]

```

## colorful substances are not tested uniformly 
```{r}
ggplot(colors.data, aes(x=total.cnt)) + 
  geom_histogram(color="black", fill="#00A86B") + 
  geom_vline(aes(xintercept=median(hitrates.sp$total.cnt)),
             color="blue", linetype="dashed", linewidth=1) +
  ggtitle('Number of Spectrophotometric Assay Endpoints per Color Chemical') +
  xlab('Number of Assay Endpoints Tested') + 
  ylab('Number of Chemicals') +
  theme_bw()

```

## create RData 
```{r, eval = FALSE}
save(mc5, 
     sc2, 
     long_hitc_final, 
     file = "./data/invitrodb_v3_5_snapshot.RData")

# save all chems in ToxCast library to get ToxPrints 
lib <- data.frame(DTXSID = hit.rates$dsstox_substance_id)
write.csv(lib, "./data/ToxCast_all_chems_invitrodb_v3_5.csv", row.names = F)

```

# What about long-chain carbons? {.tabset .tabset-fade .tabset-pills}

## load ToxPrints 
```{r}
toxprints <- fread("./data/ToxCast_invitrodb_v3_5_toxprints.csv")

length(unique(toxprints$DTXSID))

```
* 8909 out of the 9541 chemicals in ToxCast/Tox21 library can be mapped to the ToxPrint space. 

## pan-active chemotypes 

We wish to investigate whether chemicals with chemotypes found to be significantly enriched across the DIO/IYD assays may be pan-active across all protein-based/cell-free assays found in ToxCast/Tox21. 

```{r}
long_hitc_final$cell_format <- assays$cell_format[match(long_hitc_final$aeid, assays$aeid)]

# these are overlapping chem sets
c8 <- toxprints[`chain:alkaneLinear_octyl_C8` == 1, DTXSID]
c10 <- toxprints[`chain:alkaneLinear_decyl_C10` == 1, DTXSID]
c12 <- toxprints[`chain:alkaneLinear_dodedyl_C12` == 1, DTXSID]
c14 <- toxprints[`chain:alkaneLinear_tetradecyl_C14` == 1, DTXSID]
all.chains <- Reduce(union, list(c8, c10, c12, c14))

length(unique(all.chains)) 

```

* There are 391 unique chemicals found in the ToxCast/Tox21 library that had at least one of these chemotypes: 
  * chain:alkaneLinear_octyl_C8
  * chain:alkaneLinear_decyl_C10
  * chain:alkaneLinear_dodedyl_C12
  * chain:alkaneLinear_tetradecyl_C14

## compute hit rates in different sets of assays 
```{r, fig.width = 15, fig.height = 11}
tc.longchainc.data <- hit.rates[dsstox_substance_id %in% all.chains]

# describe each chem by its largest carbon length 
tc.longchainc.data[dsstox_substance_id %in% c8, carbon_length := "C8"]
tc.longchainc.data[dsstox_substance_id %in% c10, carbon_length := "C10"]
tc.longchainc.data[dsstox_substance_id %in% c12, carbon_length := "C12"]
tc.longchainc.data[dsstox_substance_id %in% c14, carbon_length := "C14"]

# hit rates across all ToxCast assays for just these carbon chains 
tc.gg <- ggplot(tc.longchainc.data, aes(x=active.pct)) + 
  geom_histogram(color="black", alpha = 0.6) + 
  geom_vline(aes(xintercept=median(hit.rates$active.pct)),
             color="red", linetype="dashed", linewidth=1) +
  annotate(geom = "text", x = median(hit.rates$active.pct) + 3, y = 90, 
           label = paste0(round(median(hit.rates$active.pct), 1), "%")) +
  ggtitle('Hit Rates across all ToxCast Assays for Long-Chain Carbons') +
  xlab('% of Assays found to be Active in') + ylab('Number of Chemicals') +
  theme_bw() +
  theme(plot.title = element_text(size = 16),
        axis.title.x = element_text(size = 16), 
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 13), 
        axis.text.y = element_text(size = 13))

# compute hit rates in only Novascreen Assays 
nvs.data <- long_hitc_final[grep("^NVS", long_hitc_final$aenm), ]

hitrates.nvs <- nvs.data[, list(
  total.cnt  = .N, #total number of aeids tested for each chem
  active.cnt  = as.double(lw(hitc==1)),  # active count
  inactive.cnt  = as.double(lw(hitc==0)),  #inactive count
  active.pct = (lw(hitc==1)/.N)*100, #active percent
  inactive.pct = (lw(hitc==0)/.N)*100 #inactive percent
), by = list(chnm, dsstox_substance_id)]

nvs.longchainc.data <- hitrates.nvs[dsstox_substance_id %in% all.chains]

nvs.gg <- ggplot(nvs.longchainc.data, aes(x=active.pct)) + 
  geom_histogram(color="black", alpha = 0.6) + 
  geom_vline(aes(xintercept=median(hitrates.nvs$active.pct)),
             color="red", linetype="dashed", linewidth=1) +
  annotate(geom = "text", x = median(hitrates.nvs$active.pct) + 2, y = 90, 
           label = paste0(round(median(hitrates.nvs$active.pct), 1), "%")) +
  ggtitle('Hit Rates across all Novascreen Assays for Long-Chain Carbons') +
  xlab('% of Assays found to be Active in') + ylab('Number of Chemicals') +
  theme_bw() +
  theme(plot.title = element_text(size = 16),
      axis.title.x = element_text(size = 16), 
      axis.title.y = element_text(size = 16),
      axis.text.x = element_text(size = 13), 
      axis.text.y = element_text(size = 13))


# compute hit rates in just the deiodinase assays
deiod.data <- long_hitc_final[grep("\\w+h(DIO|IYD)\\w+", long_hitc_final$aenm), ]

hitrates.deiod <- deiod.data[, list(
  total.cnt  = .N, #total number of aeids tested for each chem
  active.cnt  = as.double(lw(hitc==1)),  # active count
  inactive.cnt  = as.double(lw(hitc==0)),  #inactive count
  active.pct = (lw(hitc==1)/.N)*100, #active percent
  inactive.pct = (lw(hitc==0)/.N)*100 #inactive percent
), by = list(chnm, dsstox_substance_id)]

deiod.longchainc.data <- hitrates.deiod[dsstox_substance_id %in% all.chains]

deiod.gg <- ggplot(deiod.longchainc.data, aes(x=active.pct)) + 
  geom_histogram(color="black", alpha = 0.6) + 
  geom_vline(aes(xintercept=median(hitrates.deiod$active.pct)),
             color="red", linetype="dashed", linewidth=1) +
  annotate(geom = "text", x = median(hitrates.deiod$active.pct) + 3, y = 90, 
           label = paste0(round(median(hitrates.deiod$active.pct), 1), "%")) +
  ggtitle('Hit Rates across hDIO/IYD Assays for Long-Chain Carbons') +
  xlab('% of Assays found to be Active in') + ylab('Number of Chemicals') +
  theme_bw() +
  theme(plot.title = element_text(size = 16),
        axis.title.x = element_text(size = 16), 
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 13), 
        axis.text.y = element_text(size = 13))

# compute hit rates in just the cell-free assays
cellfree.data <- long_hitc_final[cell_format %in% c("cell-free", "tissue-based cell-free")]

hitrates.cf <- cellfree.data[, list(
  total.cnt  = .N, #total number of aeids tested for each chem
  active.cnt  = as.double(lw(hitc==1)),  # active count
  inactive.cnt  = as.double(lw(hitc==0)),  #inactive count
  active.pct = (lw(hitc==1)/.N)*100, #active percent
  inactive.pct = (lw(hitc==0)/.N)*100 #inactive percent
), by = list(chnm, dsstox_substance_id)]

cf.longchainc.data <- hitrates.cf[dsstox_substance_id %in% all.chains]

cf.gg <- ggplot(cf.longchainc.data, aes(x=active.pct)) + 
  geom_histogram(color="black", alpha = 0.6) + 
  geom_vline(aes(xintercept=median(hitrates.cf$active.pct)),
             color="red", linetype="dashed", linewidth=1) +
  annotate(geom = "text", x = median(hitrates.cf$active.pct) + 3, y = 90, 
           label = paste0(round(median(hitrates.cf$active.pct), 1), "%")) +
  ggtitle('Hit Rates across all Cell-free Assays for Long-Chain Carbons') +
  xlab('% of Assays found to be Active in') + ylab('Number of Chemicals') +
  theme_bw() +
  theme(plot.title = element_text(size = 16),
        axis.title.x = element_text(size = 16), 
        axis.title.y = element_text(size = 16),
        axis.text.x = element_text(size = 13), 
        axis.text.y = element_text(size = 13))

tc.gg <- tc.gg + xlab("") 
nvs.gg <- nvs.gg + xlab("") + ylab("") 
cf.gg <- cf.gg + ylab("")

ggarrange(tc.gg, nvs.gg, deiod.gg, cf.gg, 
          nrow = 2, ncol = 2)

```

* All dashed lines indicate the median hit rate among *all* chemicals tested for that set of assays (e.g. all of invitrodb, spectrophotometric assays alone, Novascreen assays, etc). 

## what are the long-chain C's that were active in all of the deiodinase assays?

Out of the 112 unique long-chain carbons tested in the hDIO/IYD assays...
```{r}
deiod.longchainc.data[active.pct == 100]

```

## save to RData file for prioritization 
```{r, eval = FALSE}
e <- new.env(parent = emptyenv())
(load('./data/Deiodinases/DIO-IYD_invitrodb_v3_5_Enrichment.RData', envir = e))
e$dio.panactive <- deiod.longchainc.data[active.pct == 100, dsstox_substance_id]
e$colors.panactive <- hit.rates.colors[active.pct >= 30, dsstox_substance_id] # promiscuous across db
do.call("save", c(ls(envir = e), list(envir = e, file = './data/Deiodinases/DIO-IYD_invitrodb_v3_5_Enrichment.RData')))

```