---
title: "Overview of Thyroid-Related Data in ToxCast"
author: "Kimberly Truong"
date: "2024-07-23"
output:
  html_document:
    code_folding: hide
    collapsed: yes
    df_print: paged
    lightbox: no
    number_sections: no
    self_contained: yes
    thumbnails: no
    toc: yes
    toc_float: yes
    keep_md: no 
  pdf_document:
    toc: yes
---

# Load Invitrodb data 
```{r setup, include = FALSE}

library(knitr)
knitr::opts_knit$set(root.dir = 'C:/Users/ktruong/OneDrive - Environmental Protection Agency (EPA)/Profile/Documents/thyroid_prioritization')

```

```{r, warning = FALSE, message = FALSE}

rm(list=ls())
library(data.table)
library(tcpl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(viridis)
library(RColorBrewer)
library(colorspace)
library(scales)
library(ggpubr)
library(pheatmap)
library(dendextend)

load('./data/invitrodb_v3_5_thyroid_data.RData')

# create a `not in` operator
`%notin%` <- Negate(`%in%`)

```

# Combining sc and mc data {.tabset .tabset-fade .tabset-pills}

## Get AC50s, hitcalls, filter out control wells
```{r test-data, message = FALSE}
mc5[hitc == 0, modl_ga := 4] # negative in mc => ac50 is assigned a 4

long_mc5 <- mc5 %>% 
  select(dsstox_substance_id,aenm,modl_ga) %>% 
  group_by(dsstox_substance_id,aenm) %>%
  summarise(modl_ga = mean(modl_ga)) %>%
  filter(!is.na(dsstox_substance_id)) # this filters out control wells 

long_sc2 <- sc2 %>% 
  select(dsstox_substance_id,aenm,hitc) %>%
  group_by(dsstox_substance_id,aenm) %>%
  summarise(hitc = max(hitc)) %>%
  filter(!is.na(dsstox_substance_id))

```

## Assign dummy values for chemicals with just sc data
```{r data-wrangling}
# get ac50 and/or sc-hitc for each chemical
long_all <- long_mc5 %>% 
  full_join(long_sc2, by = c("dsstox_substance_id","aenm")) 

# when there is only sc data, assign a 5 regardless of +/-
long_hmap <- long_all %>%
  rowwise() %>%
  mutate(var_val = case_when(
    is.na(modl_ga) ~ 5,
    TRUE ~ modl_ga
  )) %>%
  ungroup()

# pivot wider
Mac50 <-  long_hmap %>% 
  select(dsstox_substance_id, aenm, var_val) %>%
  pivot_wider(names_from = aenm, values_from = var_val) %>%
  as.data.frame() # rownames can't be set on tibble otherwise

Mac50[is.na(Mac50)] = 6 # chem not tested in given assay 
rownames(Mac50) <- Mac50$dsstox_substance_id
Mac50 <- Mac50[, -1]

head(Mac50)

```

## Form a hitcall matrix 
```{r}
long_hmap <- long_hmap %>% 
  mutate(hitc = case_when(
    modl_ga == 4 | modl_ga == 5 ~ 0, 
    modl_ga %notin% c(4,5) ~ 1
  ))

Mhc <- long_hmap %>% 
  select(dsstox_substance_id, aenm, hitc) %>% 
  pivot_wider(names_from = aenm, values_from = hitc) %>% 
  as.data.frame()
rownames(Mhc) <- Mhc$dsstox_substance_id
Mhc <- Mhc[, -1]

```

# Overview of thyroid-related data {.tabset .tabset-fade .tabset-pills}
## Assessing Total Coverage 

There are 102 unique assay endpoints in ToxCast/Tox21 that are thyroid-related. 
```{r uniq-aeids}
# how many aeids are there?
ncol(Mhc)

```

```{r}
Mtest = Mhc
Mtest[!is.na(Mhc)] = 1 
Mtest[is.na(Mhc)] = 0
cs = colSums(Mtest) # chems in each aeid
rs = rowSums(Mtest) # aeids tested for each chem

# Number of Assay Endpoints per Chem 
hist(rs, main = 'Number of Thyroid-related Endpoints for each Chemical',
     xlab = '# of Aeids per Chemical', 
     ylab = '# of Chemicals')

summary(rs)

# Number of Chems per Endpoint 
hist(cs, main = 'Number of Chemicals Tested in each Thyroid-Related Endpoint',
     xlab = '# of Chemicals Tested in an Aeid', 
     ylab = '# of Aeids')

summary(cs)

```
## ToxCast Coverage 
Removing the 27 Tox21 assays to find sets of broader chemical coverage...
```{r}
# How many Tox21 assays are there? 
ans <- ace.thyroid[grepl("TOX21", ace.thyroid$assay_component_endpoint_name), ]
tox21.aenm <- unique(ans$assay_component_endpoint_name)

# look at just the ToxCast assays 
Mhc.tc <- Mhc[, names(Mhc) %notin% tox21.aenm]
Mhc.tc <- Mhc.tc[rowSums(is.na(Mhc.tc)) != ncol(Mhc.tc), ]

Mac50.tc <- Mac50[, names(Mac50) %notin% tox21.aenm]
Mac50.tc <- Mac50.tc[rowSums(Mac50.tc == 6) != ncol(Mac50.tc), ]

Mtest.tc = Mhc.tc
Mtest.tc[!is.na(Mhc.tc)] = 1
Mtest.tc[is.na(Mhc.tc)] = 0 
cs.tc = colSums(Mtest.tc)
rs.tc = rowSums(Mtest.tc)

# Number of Assay Endpoints per Chem 
hist(rs.tc, main = 'Number of Thyroid-related ToxCast Endpoints for each Chemical',
     xlab = '# of Aeids per Chemical', 
     ylab = '# of Chemicals')

# Number of Chems per Endpoint 
hist(cs.tc, main = 'Number of Chemicals Tested in each Thyroid-Related Endpoint from ToxCast',
     xlab = '# of Chemicals in an Aeid', 
     ylab = '# of Aeids')

```

## How many hits were sc-positive but no mc?
There is a single row for each unique DTXSID x aeid pair (due to using `tcpl:: tcplSubsetChid()`). 

Table summarizes the number of chemicals tested positive in sc but were never tested in mc for whatever reason. The deiodinase assays were among the ones with the highest counts. Nchems gives the total number of unique chemicals tested in that assay. 

```{r}
setDT(long_all)
long_all[, Nchems := length(unique(dsstox_substance_id)), by = aenm]

# how many were tested to be sc == 1 but no mc?
scpos.nomc.out <- long_all[hitc == 1 & is.na(modl_ga), .(scpos_nomc = .N), by = .(aenm, Nchems)
                           ][order(-scpos_nomc)]

scpos.nomc.out

```

## analyze hDIO/IYD assays
```{r}
deiod.meta <- ace.thyroid[thyroid_related_target == 'Deiodinases', 
                          .(aeid, aenm = assay_component_endpoint_name)]

# ignore amphibian assays 
h.deiod.aeids <- deiod.meta[grepl("\\w+h(DIO|IYD)\\w+", deiod.meta$aenm), aeid]
names(h.deiod.aeids) <- c('DIO1', 'DIO2', 'DIO3', 'IYD')

# add aeid column to long_all and scpos.nomc.out 
setnames(ace.thyroid, "assay_component_endpoint_name", "aenm")
long_all$aeid <- ace.thyroid$aeid[match(long_all$aenm, ace.thyroid$aenm)]
scpos.nomc.out$aeid <- ace.thyroid$aeid[match(scpos.nomc.out$aenm, ace.thyroid$aenm)]

# get all the chems that were sc-positive but missing mc among deiod assays
deiod.no.mc <- long_all[aeid %in% h.deiod.aeids & is.na(modl_ga) & hitc == 1, 
                        .(aeid, aenm, dsstox_substance_id)]

# compare coff, max_med for each assay 
list_of_df <- vector("list", length(h.deiod.aeids))
summarize_by_assay <- function(endpoint.id) {
  chems <- deiod.no.mc[aeid == endpoint.id, dsstox_substance_id]
  out <- sc2[aeid == endpoint.id & dsstox_substance_id %in% chems, 
             .(dsstox_substance_id, coff, max_med)]
  return(summary(out))
}

list_of_df <- lapply(h.deiod.aeids, summarize_by_assay)
list_of_df

```
A 50%-inhibition hard cutoff was used to pass chemicals to mc for the DIO/IYD assays. 

# Hierarchical Clustering of Chemicals {.tabset .tabset-fade .tabset-pills}

## categorize thyroid targets 
```{r}
# categorize thyroid targets 
ace.thyroid[, category := as.character(NA)]
ace.thyroid[thyroid_related_target %in% c('TPO', 'NIS', 'Deiodinases'), category := 'TH_economy']
ace.thyroid[thyroid_related_target %in% c('TRHR', 'TSHR', 'THR'), category := 'endocrine']
ace.thyroid[thyroid_related_target %in% c('AhR', 'CAR', 'PXR'), category := 'liver']
ace.thyroid[thyroid_related_target == 'Binding', category := 'TH_binding']

```

## run hclust
```{r, fig.height = 15, fig.width = 15}
Mac50.tc.mat <- as.matrix(Mac50.tc)

# reorder assays by category
Mac50.tc.aenm <- colnames(Mac50.tc.mat)
Mac50.tc.categ <- sapply(Mac50.tc.aenm, function(k) {
  ace.thyroid[aenm == k, unique(category)]
})
metadata <- data.frame(category = Mac50.tc.categ, row.names = Mac50.tc.aenm)
metadata$category <- as.factor(metadata$category)
metadata <- metadata %>% arrange(factor(category, levels = c('endocrine', 'liver', 'TH_synthesis', 'TH_binding')))

Mac50.tc.mat <- Mac50.tc.mat[, rownames(metadata)]
hclust_chems <- hclust(dist(Mac50.tc.mat), method = 'complete')

as.dendrogram(hclust_chems) %>%
  plot(horiz = TRUE)

# choose k for hierarchical clustering 
kh <- 5
chem_cluster <- cutree(tree = as.dendrogram(hclust_chems), k = kh)
clust_dat <- as.data.frame(chem_cluster)
colnames(clust_dat) <- 'cluster'
clust_dat$cluster <- as.factor(clust_dat$cluster)

```

## ToxCast heatmap

```{r, fig.width=18,fig.height=14}
ac50_bk <- seq(-5.5, 3, by = 0.1)
bk2 <- seq(3.5, 6.5, by = 1)
bk <- c(ac50_bk, bk2)

ac50.cols <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(n = length(ac50_bk)- 1)
# ac50.cols <- viridis(n = length(ac50_bk) - 1)
# other.cols <- c('#FFFFFF', '#44BB99', '#00BAD7', '#DDDDDD')
other.cols <- c('#FFFFFF', '#66CC99', '#9999CC', '#DDDDDD')
all.cols <- c(ac50.cols, other.cols)

# change annotation colors 
annot_rcolors <- brewer.pal(n = kh + 1, name = 'Set3')[-1] # stole from Set3
annot_ccolors <- brewer.pal(n = 4, name = 'Set1')

names(annot_rcolors) <- seq_len(kh)
names(annot_ccolors) <- c('endocrine', 'liver', 'TH_binding', 'TH_economy')

ann_colors <- list(category = annot_ccolors, 
                   cluster = annot_rcolors)

tc.hmap <- pheatmap(Mac50.tc.mat,
  color = all.cols,
  breaks = bk,
  legend_breaks = c(seq(-5, 3, by = 2), 4:6),
  legend_labels = c(as.character(seq(-5,3, by = 2)), "tested negative in mc", 
                    "tested only in sc", "not tested"),
  cluster_rows = T, 
  cluster_cols = F, 
  clustering_distance_rows = 'euclidean', 
  clustering_method = 'complete', 
  # border_color = 'white', 
  # cellwidth = 20,
  # cellheight = 3,
  # cutree_rows = 5, 
  show_rownames = F, 
  annotation_col = metadata, 
  annotation_row = clust_dat, 
  annotation_colors = ann_colors, 
  annotation_legend=T)

tc.hmap

```

## clean up heatmap
```{r}
summary(cs.tc)
head(sort(cs.tc))

```

After removing these assays with few chemicals tested: 

  * ERF_ADME_hCYP2B6_dn 
  * ERF_ADME_hCYP3A4_dn
  * ERF_NR_hTHRA_Agonist
  * ATG_THRb_TRANS2_dn
  * ATG_THRb_TRANS2_up

```{r, fig.width=18,fig.height=14}
# remove assays with little chem coverage from Mac50.tc and Mhc.tc 
rmv.assays <- names(sort(cs.tc))[1:5]
jj.rmv <- which(colnames(Mhc.tc) %in% rmv.assays)

Mhc.tc.clean <- Mhc.tc[, -jj.rmv]
Mac50.tc.clean <- Mac50.tc[, -jj.rmv]

# update column (aenm) metadata
aenm.meta.clean <- metadata[(rownames(metadata) %notin% rmv.assays), , drop = FALSE]
Mac50.tc.clean <- Mac50.tc.clean[, rownames(aenm.meta.clean)]

# check to see all chems are still covered by at least one endpoint after removal
Mtest.tc.clean = Mhc.tc.clean 
Mtest.tc.clean[!is.na(Mtest.tc.clean)] = 1 
Mtest.tc.clean[is.na(Mtest.tc.clean)] = 0
rs.tc2 = rowSums(Mtest.tc.clean)
summary(rs.tc2)

# plot to double check
tc.clean.hmap <- pheatmap(Mac50.tc.clean,
  color = all.cols,
  breaks = bk,
  legend_breaks = c(seq(-5, 3, by = 2), 4:6),
  legend_labels = c(as.character(seq(-5,3, by = 2)), 
                    "tested negative in mc", "tested only in sc", "not tested"),
  cluster_rows = T, 
  cluster_cols = F, 
  clustering_distance_rows = 'euclidean', 
  clustering_method = 'complete', 
  # border_color = 'white', 
  # cellwidth = 20,
  # cellheight = 3,
  # cutree_rows = 5, 
  show_rownames = F, 
  annotation_col = aenm.meta.clean, 
  annotation_row = clust_dat, 
  annotation_colors = ann_colors, 
  annotation_legend=T)

tc.clean.hmap

ggsave(plot = tc.clean.hmap,
       units = "in",
       dpi = 300,
       width = 22, height = 14,
       device = "jpg",
       filename = "./doc/comptox-thyroid-httk/figures/supp/ToxCast_invitrodb_v3_5_thyroid_heatmap.jpg")

```

## just DIO/IYD assays

```{r, fig.width = 5, fig.height = 8}
deiod.aenm <- deiod.meta[aeid %in% h.deiod.aeids, aenm]

Mac50.deiod <- Mac50.tc.clean[, deiod.aenm]
Mhc.deiod <- Mhc.tc.clean[, deiod.aenm]
Mtest.deiod = Mhc.deiod
Mtest.deiod[!is.na(Mhc.deiod)] = 1
Mtest.deiod[is.na(Mhc.deiod)] = 0
rs.deiod = rowSums(Mtest.deiod)

Mac50.deiod <- Mac50.deiod[rs.deiod >= 1, ]
colnames(Mac50.deiod) <- c("DIO1", "DIO2", "DIO3", "IYD")

deiod.hmap <- pheatmap(Mac50.deiod,
  color = all.cols,
  breaks = bk,
  legend_breaks = c(seq(-5, 3, by = 2), 4:6),
  legend_labels = c(as.character(seq(-5,3, by = 2)), "tested negative in mc", "tested only in sc", "not tested"),
  cluster_rows = T, 
  cluster_cols = F, 
  clustering_distance_rows = 'euclidean', 
  clustering_method = 'complete', 
  show_rownames = F, 
  angle_col = 45,
  cellwidth = 40)
  # cutree_rows = 2)

ggsave(plot = deiod.hmap,
       units = "in",
       dpi = 300,
       width = 5, height = 8,
       device = "jpg",
       filename = "./doc/comptox-thyroid-httk/figures/supp/DIO-IYD_invitrodb_v3_5_heatmap.jpg")

```


