---
title: "Invitrodb_v3_5_retrieval"
author: "Katie Paul-Friedman and Kimberly Truong"
date: "4/2/2025"
output:
  html_document:
    code_folding: hide
    collapsed: yes
    df_print: paged
    lightbox: no
    number_sections: yes
    self_contained: yes
    thumbnails: no
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---
```{r setup, warning=FALSE, message=FALSE, echo=FALSE}
library(data.table)
library(tcpl)
library(dplyr)

knitr::opts_knit$set(root.dir = 'C:/Users/ktruong/OneDrive - Environmental Protection Agency (EPA)/Profile/Documents/CompTox-thyroid-httk')

```

* For invitrodb v3.5, there will be more thyroid data in there but largely for PFAS (from GLTED) compared to v3.4.

```{r, warning=FALSE, echo=FALSE}
## connect to invitrodb with tcpl
tcplConf(user='_dataminer', pass='pass', 
         db='prod_internal_invitrodb_v3_5', 
         drvr='MySQL', 
         host='ccte-mysql-res.dmap-prod.aws.epa.gov') 
tcplConfList()

```

# Load Assay annotation data {.tabset .tabset-fade .tabset-pills}

* gene level annotations help us locate the data of interest.

## Load assay information at an endpoint level

```{r, warning=FALSE, echo=FALSE}

# load assay component endpoint (ace) data
ace <- tcplQuery('SELECT * FROM assay_component_endpoint;') %>% data.table()
gene <- tcplQuery('SELECT * FROM intended_target INNER JOIN gene ON intended_target.target_id=gene.gene_id;') %>% data.table()
ace <- merge.data.table(ace,
        gene[,c('aeid', 'official_symbol','gene_name')],
                        by='aeid',
                        all.x=TRUE)

# load cytotoxicity information
cytotox <- tcplQuery("SELECT * FROM cytotox INNER JOIN chemical ON cytotox.chid=chemical.chid;") %>% data.table()

```

## Thyroid relevant assays
* based on manual review of assay endpoint IDs in the assay component endpoint table

```{r, warning=FALSE}

ace[,thyroid_related_target := as.character(NA)]

# thyroid hormone synthesis and conversion targets
ace[aeid %in% c(1508, 1509, 1824, 1848, 3090), thyroid_related_target := 'TPO']
ace[aeid %in% c(2037, 2110), thyroid_related_target := 'NIS']
ace[aeid %in% c(2309, 2532, 2533, 3032, 3091, 3092), thyroid_related_target := 'Deiodinases']

# Thyroid hormone related receptors in hypothalamus, pituitary, and thyroid (regulation)
ace[aeid %in% c(683, 2364, 2365), thyroid_related_target := 'TRHR']
ace[aeid %in% c(2040, 2043, 2046), thyroid_related_target := 'TSHR']
ace[aeid %in% c(143, 724, 803, 804, 805, 1094, 1095, 1369, 1498, 
                1499, 2226, 2227, 2230, 2236, 2237, 2240, 2241,
                2244, 2247, 2253, 2254, 2689), thyroid_related_target := 'THR']

# thyroid hormone serum binding proteins
ace[aeid %in% c(3088,3089), thyroid_related_target := 'Binding']

# liver related MIEs
ace[aeid %in% c(63, 806, 807, 1400), thyroid_related_target := 'AhR']
ace[official_symbol %in% c('Cyp1a1','Cyp1a2','CYP1A1','CYP1A2'), thyroid_related_target := 'AhR']

ace[aeid %in% c(101,
116,
712,
713,
2047,
2048,
2049,
2050), thyroid_related_target := 'CAR']
ace[official_symbol %in% c('CYP2B6'), thyroid_related_target := 'CAR']

ace[aeid %in% c(103,
135,
721,
2363,
2362), thyroid_related_target:= 'PXR']
ace[official_symbol %in% c('CYP3A4','CYP3A5','CYP3A7','Cyp3a2','Cyp3a23/Cyp3a1' ), thyroid_related_target := 'PXR']

```

```{r, warning=FALSE}

ace.thyroid <- ace[!is.na(thyroid_related_target)]

```

## Pull the quantitative data

```{r, warning=FALSE, echo=FALSE}

# load multi-conc data
# use tcplSubsetChid to downselect samples to a single representative sample per chemical considering quality of the spids 
mc5 <- tcplPrepOtpt(tcplSubsetChid(tcplLoadData(val=ace.thyroid$aeid, lvl=5, type = 'mc',fld='aeid')))
mc6 <- tcplPrepOtpt(tcplLoadData(lvl=6, fld='m4id', val=mc5$m4id, type='mc'))
setDT(mc6)
mc6_mthds <- mc6[ , .( mc6_mthd_id = paste(mc6_mthd_id, collapse=",")), by = m4id]
mc6_flags <- mc6[ , .( flag = paste(flag, collapse=";")), by = m4id]
mc5$mc6_flags <- mc6_mthds$mc6_mthd_id[match(mc5$m4id, mc6_mthds$m4id)]
mc5[, flag.length := ifelse(!is.na(mc6_flags), count.fields(textConnection(mc6_flags), sep =','), NA)]

# filter the dataset, with coarse filters
mc5[hitc==1 & flag.length < 3, use.me := 1]
mc5[hitc==1 & is.na(flag.length), use.me := 1]
mc5[hitc==1 & flag.length >= 3, use.me := 0]
mc5[fitc %in% c(36,45), use.me := 0]
mc5[hitc==-1, use.me := 0] # data could not be curve-fitted (usually due to having < 4 concs)
mc5[use.me==0, modl_ga := as.numeric(NA)] 
mc5[use.me==0, hitc := 0] 
mc5[hitc==0, modl_ga := as.numeric(NA)]

mc5[hitc==1,ac50_uM := ifelse(!is.na(modl_ga), 10^modl_ga, NA)]
mc5.hitc1 <- mc5[hitc==1]

```

## Pull the sc data

```{r, warning=FALSE}

sc2 <- tcplPrepOtpt(tcplSubsetChid(tcplLoadData(lvl=2, type='sc',val=ace.thyroid$aeid, fld='aeid'), type = 'sc')) # for max response and hitcally summary
sc1 <- tcplPrepOtpt(tcplLoadData(lvl=1,type='sc',val=ace.thyroid$aeid, fld='aeid')) # for log concentration

```

## Create RData

```{r, warning=FALSE, eval = FALSE}

save(ace,
     ace.thyroid,
     cytotox,
     gene,
     mc5,
     mc5.hitc1,
     sc1, 
     sc2, file='./data/invitrodb_v3_5_thyroid_data.RData')



```

